---
title: "Clean St. Louis Public Schools Discipline Incident"
author: "Ibuki Ogasawara"
date: "February 25, 2016"
output: html_document
---
This R markdown file cleans the datasets of discipline incidents in St. Louis public school from 2008 to 2015. Before running the code, some initial steps are required. (See footnotes.)  

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #arrange, select
```

####1. Download and Read Datasets from GitHub
```{r}
mydata1 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Discipline_Incidents/SLED_Total_Incidents.csv")
Total_Incidents <- read.csv(text = mydata1, quote = "")
mydata2 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Discipline_Incidents/Drug_Incident_Rate.csv")
Drug <- read.csv(text = mydata2, quote = "")
mydata3 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Discipline_Incidents/Violence_Rate.csv")
Violence <- read.csv(text = mydata3, quote = "")
mydata4 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Discipline_Incidents/Weapon_Indident_Rate.csv")
Weapon <- read.csv(text = mydata4, quote = "")
mydata5 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Discipline_Incidents/Out_of_School_Suspension_Rate.csv")
Suspension <- read.csv(text = mydata5, quote = "")

#Rename the columns
names(Total_Incidents) <- c("X", "School", "Year", "Enrollment", "TotalIncident")
names(Drug) <- c("School", "Year", "DrugRate")
names(Violence) <- c("School", "Year", "ViolenceRate")
names(Weapon) <- c("School", "Year", "WeaponRate")
names(Suspension) <- c("School", "Year", "SuspensionRate")
```

####3. Arrange Total Incidents Data Frame
We arrange "Total_Incidents" by "Name" alphabetically since only this dataset is ordered differently from others. 
```{r}
Total_Arranged <- arrange(Total_Incidents, School)
```

####4. Join All the Data Frames
We take "School" and "Year" column from "Total_Arrange," and for other data frames, we only take ones that are only necessary.  
```{r}
Total_INCI <- Total_Arranged[,c(2:5)]
Drug_INCI <- as.data.frame(Drug[,3])
Violence_INCI <- as.data.frame(Violence[,3])
Weapon_INCI <- as.data.frame(Weapon[,3])
Suspension_INCI <- as.data.frame(Suspension[,3])
#Check dimentions
sapply(list(Total_INCI, Drug_INCI, Violence_INCI, Weapon_INCI, Suspension_INCI), dim) 
DISC_INCI <- cbind.data.frame(Total_INCI, Drug_INCI, Violence_INCI, Weapon_INCI, Suspension_INCI)
#Rename the columns
names(DISC_INCI) <- c("School", "Year", "Enrollment", "TotalIncident",
                      "DrugRate", "ViolenceRate", "WeaponRate", "SuspensionRate")
```

####5. Assign School Number
We will assign school number to each school by extracting the nubmers in the parentheses in the "School" column. 
```{r}
#Function that extracts the 6th last character through 3rd last character of characters. 
substrRight_2 <- function(x){
  substr(x, nchar(x)-5, nchar(x)-2)
}
#Apply the function and convert the output into numeric
DISC_INCI$SchNum <- as.numeric(substrRight_2(as.character(DISC_INCI$School)))
```

####6. Delete Missing Rows
Now, we delete the row with all "NA"s. They are missing either because the school had been closed or had not yet been opened.
```{r}
#Check if each column for sure has the same number of non-NA cells
sapply(list(table(DISC_INCI$TotalIncident), table(DISC_INCI$Year), table(DISC_INCI$Enrollment), table(DISC_INCI$DrugRate), table(DISC_INCI$ViolenceRate), table(DISC_INCI$WeaponRate), table(DISC_INCI$SuspensionRate)), sum)

DISC_Reduced <- DISC_INCI[complete.cases(DISC_INCI),]

sapply(list(table(DISC_Reduced$TotalIncident), table(DISC_Reduced$Year), table(DISC_Reduced$Enrollment), table(DISC_Reduced$DrugRate), table(DISC_Reduced$ViolenceRate), table(DISC_Reduced$WeaponRate), table(DISC_Reduced$SuspensionRate)), sum)

#One of the rows has been missed because it has NAs in it even though the total number of incident is indicated.
#We add the row: 456th "LONG MIDDLE COMMUNITY ED. CTR. (3260)" 2015 column into where it is supposed to be. 
missed_row <- DISC_INCI[456, ]
DISC_Complete <- rbind(DISC_Reduced, missed_row)
```

####7. Compute Number of Incidents
Since four of the columns are the incidnet rates (percentage), we compute the number of occurance based on "Enrollment". 
```{r}
DISC_Complete$DrugNum <- DISC_Complete$DrugRate*DISC_Complete$Enrollment/100
DISC_Complete$ViolenceNum <- DISC_Complete$ViolenceRate*DISC_Complete$Enrollment/100
DISC_Complete$WeaponNum <- DISC_Complete$WeaponRate*DISC_Complete$Enrollment/100
DISC_Complete$SuspensionNum <- DISC_Complete$SuspensionRate*DISC_Complete$Enrollment/100
```

####8. Select Columns
Finally, we will extract the columns that we need and arrange it by the "SchNum"  
We do not use "ViolenceNum" and "WeaponNum" since there are so many missing cells. 
```{r}
DISC_INCI_Select <- dplyr::select(DISC_Complete, SchNum, Year, TotalIncident, DrugNum, SuspensionNum)
DISC_INCI_Final <- arrange(DISC_INCI_Select, SchNum)
```
####9. Save the data frame
```{r}
write.csv(DISC_INCI_Final, "SLED_Discipline_Incident_Cleaned.csv")
```

#### Footnote:
The original data source: Missouri Department of Elementary & Secondary Education   
http://mcds.dese.mo.gov/quickfacts/SitePages/DistrictInfo.aspx?ID=__bk8100130013005300130013005300  

Click under "ST. LOUIS CITY -- Performance Data", "Discipline Incidents" - "Building"  

Initial cleaning process:  
1. In Excel, deleted commas in one the school names (EARL NANCE, SR. ELEM. (5610))  
2. In Excel, deleted columns that indicate the graduation rate by race   
3. Import the "Total Discipline Incident" dataset into Rstudio and deleted commas in the column "ENROLLMENT" by
```{r eval=FALSE}
Enroll <- Total_Incidents$ENROLLMENT
enroll.cha <- as.character(Enroll)
enroll.num <- as.numeric(gsub(",", "", enroll.cha))
Total_Incidents$ENROLLMENT <- enroll.num
write.csv(Total_Incidents, "SLED_Total_Incidents.csv")
```







