---
title: "Clean St. Louis Public School Information Shapefile"
author: "Ibuki Ogasawara"
date: "February 29, 2016"
output: html_document
---
This R markdown file cleans and transforms the point shapefile of neighborhoods in St. Louis. 

####0. Install Required Packages
```{r message=FALSE}
if(!require(maptools))install.packages("maptools")
library(maptools) #readShapeSpatial (read shapefile)
if(!require(rgdal))install.packages("rgdal")
library(rgdal) #spTransform (transforms the coordinate system)
```

####1. Download and Check the Properties of the Shapefile
```{r message=FALSE}
#The shapefile is downloaded from Data.gov (https://catalog.data.gov/dataset/schools-k-12-mo-2012-public-schools-shp). 
zipurl <- "ftp://msdis.missouri.edu/pub/Facilities_Structures/MO_2012_Public_Schools_shp.zip"

#The function is take from the following website: https://dzone.com/articles/function-batch-downloading
url_shp_to_spdf <- function(URL) {
 
  require(rgdal)
 
  wd <- getwd()
  td <- tempdir()
  setwd(td)
 
  temp <- tempfile(fileext = ".zip")
  download.file(URL, temp)
  unzip(temp)
 
  shp <- dir(tempdir(), "*.shp$")
  lyr <- sub(".shp$", "", shp)
  y <- lapply(X = lyr, FUN = function(x) readOGR(dsn=shp, layer=lyr))
  names(y) <- lyr
 
  unlink(dir(td))
  setwd(wd)
  return(y)
}
#Run the funciton with the "zipurl"
y <- url_shp_to_spdf(zipurl)
z <- unlist(unlist(y))
#Define the shapefile in the output. 
MO_Sch <- z$MO_2012_Public_Schools_shp
#subset the shapefile by "CtyDist" so that we only have the points in the city of St. Louis
SLC_sch <- MO_Sch[MO_Sch$CtyDist=="115115",]
#Check what fields that this shapefile has. 
names(SLC_sch)
#Check the key field of joining dataset. 
SLC_sch$SchNum
#Check the original length of the shapefile
length(SLC_sch)
```

####2. Select Attributes
First, remove columns that we do not need and keep only "SchNum" and the names of "Facility"
```{r}
SL_SCH_Reduced <- SLC_sch[, -c(1,3,c(5:23))]
```

####3.0. Add Points - Identify Schools That Need to Be Added - 
Now, we will start adding spacial points to the shapefile because it does not include the schools that have been closed before 2012 or has been established after 2012. 
```{r message=FALSE}
#First, we need to identify which points are need to be added. 
SLED_Demographic <- read.csv("/data/home/graphics/MAP_16_Spring/Final/4. Initial Clean Data/SLED_Demographic_Cleaned.csv")
```

####3.1. Add Points - Create a List of Missing Schools - 
We create a list of schools that DO NOT consistently appear in the St. Louis public schools demographic dataset through out the years (2008-2015) because they are the ones that are likely not to be in the shapefile. 
```{r results=FALSE}
table(SLED_Demographic$SCHOOL_NAME)
```
Now, we search for UTM coordinate for the schools that are missing.   
Search for the info using  http://www.publicschoolreview.com/ and "google map" and  
http://www.mappingsupport.com/p/recreation/utm_coordinates_topo_map.html and http://www.geoplaner.com/

ACAD OF ENVT SCI/MATH ELEM (4990):exists in shapefile
ACAD OF ENVT SCI/MATH MIDDLE (3250):exists in shapefile
BADEN ELEM. (4080):closed 2010 738854.1/4287327
BIG PICTURE ELEM. ALT. (1021): closed 2010 735553.1/4281711
BIG PICTURE MIDDLE & HIGH @ NW (1023): closed 2012 740434.1/4278198
BIG PICTURE MIDDLE ALT. (1022): closed 2010 740520.1/4282681
BLEWETT MIDDLE (3020): closed 2010 743390.1/4280960
BLOW MIDDLE (3040): closed 2009 739030.1/4270954
BUNCHE INTERNATIONAL STUDIES (3110): New? 744161.1/4277931
CLARK ELEM. (4320): New? Outside of the city area.
COLLEGIATE SCHOOL OF MED/BIO (1510):exists in shapefile
GALLAUDET SCH. FOR DEAF ELEM. (4720): closed 2012 740432.1/4278051
GRISCOM JUVENILE DETENTION CTR (1015): It exists now 740595.1/4280936
GUNDLACH ELEM. (4760): closed 2009 737940.1/4284457
HUMBOLDT ACAD OF HIGHER LRNING (4960):exists in shapefile
L'OUVERTURE MIDDLE (3280):exists in shapefile
MARK TWAIN ELEM. (5280): New? Outside of the city area.
MCKINLEY/CLASSICAL JR. ACAD. (3130): Same building as MCLA 742120.1/4277159
MITCHELL ELEM. (5540): closed 2009 Same building as KIPP Victory Academy? 736694.1/4282550
NORTHWEST MIDDLE (3310): closed 2010 Same building as NW Acad of Law 738651.1/4287176
ROOSEVELT 9TH GRADE CTR. (1640):closed 2010 743188.1/4276215
SCRUGGS ELEM. (5740): closed 2010 739662.1/4273410
SHEPARD ELEM. (5820): closed 2010 742985.1/4275099
SHERMAN ELEM. COMM. ED. CTR. (5840):exists in shapefile
SIMMONS ELEM. (5880): closed 2010 740321.1/4282981
STOWE PREPARATORY JR. HIGH (2070): closed 2010 736845.1/4284089
THE COLLEGE PREPARATORY HIGH (1550):exists in shapefile
VASHON 9TH GRADE CTR. (1600): closed 2009 740941.1/4282162
WYMAN ELEM. (6140): Unsure, New? 740604.1/4278183

####3.2. Add Points - Add Table and Coordinates - 
We have to add the table and spacial information separately. 
```{r}
#create additional rows for "SL_SCH_Reduced@data". 
Additional_Schools <- as.data.frame(rbind(c(4080, "Baden Elem"),
                                          c(1021, "Big Picture Elem Alt"),
                                          c(1023, "Big Picture Middle & High"),
                                          c(1022, "Big Picture Middle Alt"),
                                          c(3020, "Blewett Middle"),
                                          c(3040, "Blow Middle"),
                                          c(3110, "Bunche Intl Studies"),
                                          c(4720, "Gallaudet Sch for Deaf Elem"),
                                          c(1015, "Griscom Juvenile Detention CTR"),
                                          c(4760, "Gundlach Elem"),
                                          c(3130, "McKinley/Classical JR. Acad"),
                                          c(5540, "Mitchell Elem"),
                                          c(3310, "Northwest Middle"),
                                          c(1640, "Roosevelt 9th Grade CTR"),
                                          c(5740, "Scruggs Elem"),
                                          c(5820, "Shepard Elem"),
                                          c(5880, "Simmons Elem"),
                                          c(2070, "Stowe Preparatory JR. High"),
                                          c(1600, "Vashon 9th Grade CTR"),
                                          c(6140, "Wyman Elem")))

#Name the columns same as the ones in the shapefile. 
names(Additional_Schools) <- c("SchNum", "Facility")

#create additional rows for "SL_SCH_Reduced@coords"
Additional_Points <- as.matrix(rbind(as.numeric(c(738854.1,4287327)),
                                      as.numeric(c(735553.1,4281711)),
                                      as.numeric(c(740434.1,4278198)),
                                      as.numeric(c(740520.1,4282681)),
                                      as.numeric(c(743390.1,4280960)),
                                      as.numeric(c(739030.1,4270954)),
                                      as.numeric(c(744161.1,4277931)),
                                      as.numeric(c(740432.1,4278051)),
                                      as.numeric(c(740595.1,4280936)),
                                      as.numeric(c(737940.1,4284457)),
                                      as.numeric(c(742120.1,4277159)),
                                      as.numeric(c(736694.1,4282550)),
                                      as.numeric(c(738651.1,4287176)),
                                      as.numeric(c(743188.1,4276215)),
                                      as.numeric(c(739662.1,4273410)),
                                      as.numeric(c(742985.1,4275099)),
                                      as.numeric(c(740321.1,4282981)),
                                      as.numeric(c(736845.1,4284089)),
                                      as.numeric(c(740941.1,4282162)),
                                      as.numeric(c(740604.1,4278183))))
#Add the coordinate matrix to the existed spacial information.
SL_SCH_Reduced@coords <- rbind(SL_SCH_Reduced@coords, Additional_Points)
#Add the dataframe to the attribute table.
SL_SCH_Reduced@data <- rbind(SL_SCH_Reduced@data, Additional_Schools)
#Check the result. 
tail(SL_SCH_Reduced@data)
#"SL_SCH_Full" has all the school points possible from 2008 to 2015. 
SL_SCH_Full <- SL_SCH_Reduced
#Check the points that are added. 
plot(SLC_sch)
plot(SL_SCH_Full)
```

####4. Reproject and Transform the Shapefile
```{r}
#Open the projection folder in the shapefile (.prj) and find the projection code in the following website:
# http://downloads.esri.com/support/documentation/ims_/Support_files/elements/pcs.htm
#In this case, it is 26915 (UTM)
proj4string(SL_SCH_Full) <- CRS("+init=esri:26915")
#epsg:4326 is for latitude and longitude coordinate system.
SL_SCH_Full_LatLong <- spTransform(SL_SCH_Full, CRS("+init=epsg:4326"))

#We want "SchNum" column to be numeric for later use.
SL_SCH_Full_LatLong@data$SchNum <- as.numeric(SL_SCH_Full_Latlong@data$SchNum)
```

####5. Save the Point Shapefile
```{r}
#Create a folder named "SL_SCH_2008_Full_shp" in the working directry 
writeSpatialShape(SL_SCH_Full_LatLong, "SL_School_Shapefile_Cleaned")
```


#### Footnote:

The St. Louis Public School Point Shapefile is obtained from DATA.GOV:  
https://catalog.data.gov/dataset/schools-k-12-mo-2012-public-schools-shp

The function that allows us to directly download the shapefile from the website is copied from the following website:  
https://dzone.com/articles/function-batch-downloading




