---
title: "Clean St. Louis Police Department Crime Data"
author: "Ibuki Ogasawara"
date: "February 8, 2016"
output: html_document
---

This R markdown file cleans the datasets of crime record of St. Louis City from 2008 to 2015. Before running the code, some initial steps are required. (See footnotes.) 

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #summarize, arrange, select, filter
```

####1. Download, Read, and Rename Datasets from GitHub
```{r message=FALSE}
mydata15 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2015.csv")
St_Louis_2015 <- read.csv(text = mydata15, quote = "")

mydata14 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2014.csv")
St_Louis_2014 <- read.csv(text = mydata14, quote = "")

mydata13 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2013.csv")
St_Louis_2013 <- read.csv(text = mydata13, quote = "")

mydata12 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2012.csv")
St_Louis_2012 <- read.csv(text = mydata12, quote = "")

mydata11 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2011.csv")
St_Louis_2011 <- read.csv(text = mydata11, quote = "")

mydata10 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2010.csv")
St_Louis_2010 <- read.csv(text = mydata10, quote = "")

mydata9 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2009.csv")
St_Louis_2009 <- read.csv(text = mydata9, quote = "")

mydata8 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/St.%20Louis%20Crime%20Data/SLPD2008.csv")
St_Louis_2008 <- read.csv(text = mydata8, quote = "")

#Rename the columns
names(St_Louis_2015) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2014) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2013) <- c("Year", "DateOccur", "Count", "Crime", "District",
                          "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2012) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2011) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2010) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2009) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")
names(St_Louis_2008) <- c("Year", "DateOccur", "Count", "Crime",
                          "District", "Neighborhood", "XCoord", "YCoord")

#Check the dimention of each table. 
sapply(list(St_Louis_2015, St_Louis_2014, St_Louis_2013, St_Louis_2012, St_Louis_2011, St_Louis_2010, St_Louis_2009, St_Louis_2008), dim)

#Create a year column to each table.
St_Louis_2015$Year <- rep(2015, 51235)
St_Louis_2014$Year <- rep(2014, 47876)
St_Louis_2013$Year <- rep(2013, 51970)
St_Louis_2012$Year <- rep(2012, 53387)
St_Louis_2011$Year <- rep(2011, 55247)
St_Louis_2010$Year <- rep(2010, 57829)
St_Louis_2009$Year <- rep(2009, 63046)
St_Louis_2008$Year <- rep(2008, 68410)
```

####2. Join the Data Frames
Now, we will combine the datasets into one dataset. 
```{r}
St_Louis <- rbind.data.frame(St_Louis_2015, St_Louis_2014, St_Louis_2013, St_Louis_2012, St_Louis_2011, St_Louis_2010, St_Louis_2009, St_Louis_2008)
```

####3. Delete Non-Criminal Incidents
When the "count" is -1, that means that the original crime reported for that complaint number was NOT indeed a crime. Therefore, we delete the rows with "count"=-1.
```{r message=FALSE}
table(St_Louis$Count)
St_Louis_Filtered <- filter(St_Louis, St_Louis$Count == 1)
```

####4. Create Crime ID
Now, we sort the dataset by the crime type and create a column called "crimeID" to simplify the column, "crime".   
There are 24 crime groups, and they are:  
1: Homicide / 2: Rape / 3: Robbery / 4: Aggravated Assault / 5: Burglary / 6: Larceny / 7: Vehicle Theft / 8: Arson / 9: Simple Assault / 10: Forgery / 11: Fraud / 12: Embezzlement / 13: Stolen Property / 14: Destruction of Property / 15: Weapons Offense / 16: Prostitution / 17: Sex Offense / 18: VMCSL / 19: Offense Against Family / 20: DWI,DUI / 21: Liquor Laws / 22: Disorderly Conduct / 23: Loitering, Begging / 24: Other Crimes.
```{r}
St_Louis_Filtered$crimeID[St_Louis_Filtered$Crime < 20000] <- 1
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 20000)
                          &  (St_Louis_Filtered$Crime < 30000)] <- 2
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 30000)
                          &  (St_Louis_Filtered$Crime < 40000)] <- 3
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 40000)
                          &  (St_Louis_Filtered$Crime < 50000)] <- 4
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 50000)
                          &  (St_Louis_Filtered$Crime < 60000)] <- 5
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 60000)
                          &  (St_Louis_Filtered$Crime < 70000)] <- 6
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 70000)
                          &  (St_Louis_Filtered$Crime < 80000)] <- 7
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 80000)
                          &  (St_Louis_Filtered$Crime < 90000)] <- 8
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 90000)
                          &  (St_Louis_Filtered$Crime < 100000)] <- 9
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 100000)
                          &  (St_Louis_Filtered$Crime < 110000)] <- 10
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 110000)
                          &  (St_Louis_Filtered$Crime < 120000)] <- 11
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 120000)
                          &  (St_Louis_Filtered$Crime < 130000)] <- 12
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 130000)
                          &  (St_Louis_Filtered$Crime < 140000)] <- 13
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 140000)
                          &  (St_Louis_Filtered$Crime < 150000)] <- 14
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 150000)
                          &  (St_Louis_Filtered$Crime < 160000)] <- 15
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 160000)
                          &  (St_Louis_Filtered$Crime < 170000)] <- 16
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 170000)
                          &  (St_Louis_Filtered$Crime < 180000)] <- 17
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 180000)
                          &  (St_Louis_Filtered$Crime < 190000)] <- 18
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 200000)
                          &  (St_Louis_Filtered$Crime < 210000)] <- 19 #(20)
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 210000)
                          &  (St_Louis_Filtered$Crime < 220000)] <- 20 #(21)
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 220000)
                          &  (St_Louis_Filtered$Crime < 230000)] <- 21 #(22)
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 240000)
                          &  (St_Louis_Filtered$Crime < 250000)] <- 22 #(24)
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 250000)
                          &  (St_Louis_Filtered$Crime < 260000)] <- 23 #(25)
St_Louis_Filtered$crimeID[(St_Louis_Filtered$Crime >= 260000)
                          |  ((St_Louis_Filtered$Crime >=190000)
                              & (St_Louis_Filtered$Crime < 200000))
                          | ((St_Louis_Filtered$Crime >=230000)
                             &(St_Louis_Filtered$Crime < 240000))] <- 24 #(26)
```

####5. Consolidate Crime Types
Consolidate the crime types into 7 different category:  
1: Homicide (1: Homicide)  
2: Assault (4: Aggravated Assault, 9:Simple Assault, 19:Offense against family)  
3: Sex Offense (2:Rape, 17:Sex Offense)  
4: Contraband (15: Weapons Offense, 16: Prostitution, 18: VMCSL, 21: Liquor Laws)  
5: Larceny (3:Robbery, 5:Burglary, 6:Larceny, 7:Vehicle Theft, 13:Stolen Property)  
6: Vandalism (8:Arson, 14:Destruction of Property)  
7: Other (20:DWI, DUI, 10: Forgery, 11: Fraud, 12: Embezzlement, 22: Disorderly Conduct, 23: Loitering, Begging, 24: Other Crimes)
```{r}
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==1] <- "Homicide"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==4
                            | St_Louis_Filtered$crimeID==9
                            | St_Louis_Filtered$crimeID==19] <- "Assault"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==2
                            | St_Louis_Filtered$crimeID==17] <- "Sex Offense"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==15
                            | St_Louis_Filtered$crimeID==16
                            | St_Louis_Filtered$crimeID==18
                            | St_Louis_Filtered$crimeID==21] <- "Contraband"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==3
                            | St_Louis_Filtered$crimeID==5
                            | St_Louis_Filtered$crimeID==6
                            | St_Louis_Filtered$crimeID==7
                            | St_Louis_Filtered$crimeID==13] <- "Larceny"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==8
                            | St_Louis_Filtered$crimeID==14] <- "Vandalism"
St_Louis_Filtered$CrimeType[St_Louis_Filtered$crimeID==10
                            | St_Louis_Filtered$crimeID==11
                            | St_Louis_Filtered$crimeID==12
                            | St_Louis_Filtered$crimeID==20
                            | St_Louis_Filtered$crimeID==22
                            | St_Louis_Filtered$crimeID==23
                            | St_Louis_Filtered$crimeID==24] <- "Other Crimes"
```

####5. Create Binary Columns for Crime Types
Create 6 other columns of dummy variables with crime types.
```{r}
St_Louis_Dummies <- mutate(St_Louis_Filtered, 
                    Homicide = as.numeric(St_Louis_Filtered$CrimeType=="Homicide"), 
                    Assault = as.numeric(St_Louis_Filtered$CrimeType=="Assault"),
                    Sex_Offense = as.numeric(St_Louis_Filtered$CrimeType=="Sex Offense"),
                    Contraband = as.numeric(St_Louis_Filtered$CrimeType=="Contraband"),
                    Larceny = as.numeric(St_Louis_Filtered$CrimeType=="Larceny"),
                    Vandalism = as.numeric(St_Louis_Filtered$CrimeType=="Vandalism"),
                    Others = as.numeric(St_Louis_Filtered$CrimeType=="Other Crimes"))

```

####6. Count the Number of Crimes
Count the number of each crime type in each neighborhood. 
```{r}
#Columns we want to group by
grp_cols <- names(St_Louis_Dummies)[c(1, 6)] #Select "Year" and "District"
#Convert character vector to list of symbols
dots <- lapply(grp_cols, as.symbol)

Total_counts <- St_Louis_Dummies %>%
        group_by_(.dots=dots) %>%
        summarise("CrimeType" = length(CrimeType))
Homi_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Homicide" = sum(Homicide))
Asl_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Assault" = sum(Assault))
Sex_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Sex_Offense" = sum(Sex_Offense))
Cont_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Contraband" = sum(Contraband))
Larc_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Larceny" = sum(Larceny))
Vand_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Vandalism" = sum(Vandalism))
Oth_counts <- St_Louis_Dummies %>% 
        group_by_(.dots=dots) %>%
        summarise("Other Crimes" = sum(Others))
```

####7. Join the Resutls
Create data frame with the counts of each crime type
```{r}
St_Louis_Cleaned <- data.frame("Year"=Total_counts[,1],
                                    "NHD_NUM"=Total_counts[,2],
                                    "Homicide"=Homi_counts[,3],
                                    "Assault"=Asl_counts[,3],
                                    "Sex Offense"=Sex_counts[,3],
                                    "Contraband"=Cont_counts[,3],
                                    "Larceny"=Larc_counts[,3],
                                    "Vandalism"=Vand_counts[,3],
                                    "Other_Crimes"=Oth_counts[,3],
                                    "Total"=Total_counts[,3])
names(St_Louis_Cleaned) <- c("Year","NHD_NUM","Homicide","Assault","SexOffense","Contraband","Larceny","Vandalism","Others","Total" )
```

####8. Save the Data Frame
```{r}
write.csv(St_Louis_Cleaned, "SLPD_Cleaned.csv")
```

#### Footnote:

Data Source: St. Louis Metropolitan Police Department  
http://www.slmpd.org/Crimereports.shtml

Initial Clean:  
The datasets were separated by month and year. We combined them to each year. 
In Excel, we also removed the columns that we do not need for our analysis. 

They are:  
X, Complaint, CodedMonth, FlagCrime, FlagUnfounded, FlagAdministrative, FlagCleanup, ILEADSAddress, ILEADSStreet, LocationName, LocationComment, CADAddress, CADStreet.