---
title: "Create NYC Data for app"
author: "Peiyun Zhang"
date: "May 6, 2016"
output: html_document
---


This R markdown file will introduce the initial steps to create an interactive app and some heat maps color by attributes.

First, we will merge the dataset into the shapefile
####1A. Read the Precinct Polygon Shapefile
```{r}
if(!require(raster))install.packages("raster")
library("raster") #merge (join shapefile and csv files)
if(!require(maptools)) install.packages("maptools")
library(maptools) #readShapeSpatial
if(!require(dplyr))install.packages("dplyr")
library("dplyr")

NYC_PCT <- readShapeSpatial("/data/home/ogasawar17/graphics/MAP_16_Spring/Final/7. Clean Data 2015/New York/NYPD_Precinct_Shapefile/nypp")
#Check the properties of the shapefile
length(NYC_PCT)
class(NYC_PCT)
proj4string(NYC_PCT)
names(NYC_PCT@data)
#check the class of the key field (NHD_NUM)
class(NYC_PCT@data$Precinct)
```

####2A. Read and Join the Attribute Dataset
Read the dataframe that we are going to merge into the shapefile
```{r}
NYC_Census <- read.csv("/data/home/ogasawar17/graphics/MAP_16_Spring/Final/7. Clean Data 2015/New York/NYC_Census_Data_by_Precinct_2010.csv")

names(NYC_Census)
#We do not need "X.1" and "X" columns
NYC_Census_Reduced <- NYC_Census[,-c(1,24)]
names(NYC_Census_Reduced)

NYC_Arrest <- read.csv("/data/home/ogasawar17/graphics/MAP_16_Spring/Final/7. Clean Data 2015/New York/NYPD_Arrest_2015.csv")

names(NYC_Arrest)
#We do not need "X" column
NYC_Arrest_Reduced <- NYC_Arrest[,-1]
names(NYC_Arrest_Reduced)


NYC_Census_Arrest <- full_join(NYC_Census_Reduced, NYC_Arrest_Reduced, by = 'Precinct')
```

####3A. Choose Variables of Our Interests
```{r}
#In our example, we will choose "Population", "White", "Black", "Hispanic", "Unplyrate", and "PvtyUnplyrate" from NYC_Census_Reduced.
#From NYC_Arrest_Reduced, we will choose "Arrest15", "Frisk15", "Search15", "Age", "Black", "White", "Asslt15", and "Homici15". 
column0 <- NYC_Census_Arrest$Precinct

column1 <- NYC_Census_Arrest$Population
column2 <- NYC_Census_Arrest$White
column3 <- NYC_Census_Arrest$Black
column4 <- NYC_Census_Arrest$Hispanic
column5 <- NYC_Census_Arrest$Unplyrate
column6 <- NYC_Census_Arrest$PvtyUnplyrate

column7 <- NYC_Census_Arrest$Arrest15
column8 <- NYC_Census_Arrest$Frisk15
column9 <- NYC_Census_Arrest$Search15
column10 <- NYC_Census_Arrest$Age15
column11 <- NYC_Census_Arrest$Black15
column12 <- NYC_Census_Arrest$White15
column13 <- NYC_Census_Arrest$Asslt15
column14 <- NYC_Census_Arrest$Homici15

NYC_Census_Arrest_Selected <- cbind.data.frame(column0, column1, column2, column3, column4, column5, column6, column7, column8, column9, column10, column11, column12, column13, column14)

names(NYC_Census_Arrest_Selected) <- c("Precinct","Popultn","White","Black","Hispnc","UnplyR","PvUnplyR","Arrest15","Frisk15","Search15","Age15","Black15","White15","Asslt15","Homici15")
```

####4A. Merge and Write a New Shapefile
```{r}
NYC_PCT_Shape_Attribute <- merge(NYC_PCT, NYC_Census_Arrest_Selected, by = 'Precinct')
#Check the result
View(NYC_PCT_Shape_Attribute@data)

proj4string(NYC_PCT_Shape_Attribute) <- CRS("+init=epsg:4326")

writeSpatialShape(NYC_PCT_Shape_Attribute, "NYC_PCT_Shapefile_Selected")
```


####1B. Read the School Point Shapefile
```{r}
NYC_SCH <- readShapeSpatial ("/data/home/ogasawar17/graphics/MAP_16_Spring/Final/7. Clean Data 2015/New York/NYC_SCH_Shapefile/NYC_SCH_Shapefile")

#Check the properties of the shapefile
length(NYC_SCH)
class(NYC_SCH)
proj4string(NYC_SCH)
names(NYC_SCH@data)

#check the class of the key field (NYC_NUM)
class(NYC_SCH@data$DBN)
```

####2B. Read the Attribute Dataset
Read the dataframe that we are going to merge into the shapefile
```{r}
School_Data <- read.csv("/data/home/ogasawar17/graphics/MAP_16_Spring/Final/7. Clean Data 2015/New York/NYC_School_Data_2015.csv")

names(School_Data)

#We do not need "X" column
School_Data_Reduced <- School_Data[,-1]
names(School_Data_Reduced)

class(School_Data_Reduced$DBN)
```

####3B. Choose Variables of Our Interests
```{r}
#In our example, we will choose "Total", "Black", "White", "Poverty", "MSELA", "MSMAT", "L34NmELA", "L34NmMAT", "TtlGrad4", and "DropOut4" from School_Data_Reduced.
 
column0 <- School_Data_Reduced$DBN

column1 <- School_Data_Reduced$Total
column2 <- School_Data_Reduced$Black
column3 <- School_Data_Reduced$White
column4 <- School_Data_Reduced$Poverty
column5 <- School_Data_Reduced$MSELA
column6 <- School_Data_Reduced$MSMAT
column7 <- School_Data_Reduced$L34NmELA
column8 <- School_Data_Reduced$L34NmMAT
column9 <- School_Data_Reduced$TtlGrad4
column10 <- School_Data_Reduced$DropOut4

School_Selected <- cbind.data.frame(column0, column1, column2, column3, column4, column5, column6, column7, column8, column9, column10)

names(School_Selected) <- c("DBN","Total","Black","White","Poverty","MSELA","MSMAT","L34NmELA","L34NmMAT","TtlGrad4","DropOut4")
```

####3B. Merge and Write a New Shapefile
```{r}
NYC_SCH_Shape_Attribute <- merge(NYC_SCH, School_Selected, by='DBN')

#The following count column is for counting the number of school in a precinct later on. 
NYC_SCH_Shape_Attribute@data$counts <- rep(1, length(NYC_SCH_Shape_Attribute))

#Check the result
View(NYC_SCH_Shape_Attribute@data)

proj4string(NYC_SCH_Shape_Attribute) <- CRS("+init=epsg:4326")

writeSpatialShape(NYC_SCH_Shape_Attribute, "NYC_SCH_Shapefile_Selected")
```