---
title: "Clean New York City Public Schools Demographics"
author: "Ibuki Ogasawara"
date: "March 16, 2016"
output: html_document
---
This R markdown file cleans the datasets of demographics of New York City public schools from 2011 to 2015. Before running the code, some initial steps are required. (See footnotes.) 

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #summarize, arrange, select, filter
```

####1. Download and Read Datasets from GitHub
```{r message=FALSE}
mydata <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYED%20Demographics/NYED_Demographics.csv")
NYED_DEMO <- read.csv(text = mydata, quote = "")
dim(NYED_DEMO)
```

####2. Reduce Columns
First, we will delete columns that we are not interested in. 
```{r message=FALSE}
names(NYED_DEMO)
NYED_DEMO_Reduced <- select(NYED_DEMO, DBN, School.Name, Year,
                            Total.Enrollment, X..Female, X..Male,
                            X..Asian, X..Black, X..Hispanic, X..White, 
                            X..Other, X..Students.with.Disabilities, 
                            X..English.Language.Learners, X..Poverty)
#rename columns
names(NYED_DEMO_Reduced) <- c("DBN", "School", "Year", 
                              "Total", "Female", "Male", 
                              "Asian", "Black", "Hispanic", "White", 
                              "Other", "SWD", "ELL", "Poverty")
```

####3. Clean "Year" Column
Now, clean the "Year" column so that it has the form 20XX instead of 20XX-XX. 
```{r}
table(NYED_DEMO_Reduced$Year)
NYED_DEMO_Reduced$Year <- substr(NYED_DEMO_Reduced$Year, 1, 4)
```

####4. Check the Class of Variables
Check if we need to change the class of the columns. In this case, all the columns that we want them to be numeric are "integer".
```{r}
list_of_names <- list(NYED_DEMO_Reduced$DBN, NYED_DEMO_Reduced$School, NYED_DEMO_Reduced$Year, NYED_DEMO_Reduced$Total, NYED_DEMO_Reduced$Female, NYED_DEMO_Reduced$Male, NYED_DEMO_Reduced$Asian, NYED_DEMO_Reduced$Black, NYED_DEMO_Reduced$Hispanic, NYED_DEMO_Reduced$White, NYED_DEMO_Reduced$Other, NYED_DEMO_Reduced$SWD, NYED_DEMO_Reduced$ELL, NYED_DEMO_Reduced$Poverty)
sapply(list_of_names, class)
```

####5. Subset by Year
Lastly, we will subset them by "Year"
```{r}
NYED_DEMO_11 <- filter(NYED_DEMO_Reduced, Year=="2011")
NYED_DEMO_12 <- filter(NYED_DEMO_Reduced, Year=="2012")
NYED_DEMO_13 <- filter(NYED_DEMO_Reduced, Year=="2013")
NYED_DEMO_14 <- filter(NYED_DEMO_Reduced, Year=="2014")
NYED_DEMO_15 <- filter(NYED_DEMO_Reduced, Year=="2015")
```

####6. Save the Data Frames
```{r}
write.csv(NYED_DEMO_15, "NYED_Demographics_2015_Cleaned.csv")
```

#### Footnote:

Data Source:NYC Department of Education  
http://schools.nyc.gov/AboutUs/schools/data/default.htm  

Under "Population and Demographics", click "Demographic Snapshot"  

Initial cleaning process:  
1. Choose the spread sheet for "school" and export it as a csv file.  
2. Deleted commas in one the school names in csv file. 