---
title: "2010 Censusdata Cleaning"
author: "Peiyun Zhang"
date: "February 4, 2016"
output: html_document
---

####0. Install required packages;
```{r,message=FALSE}
#"dplyr" package is used to do data cleaning and scraping
if(!require(dplyr))
    install.packages("dplyr")
    require("dplyr")
#"RCurl" package is used to read raw data from github
if(!require(RCurl))
    install.packages("RCurl")
    require("RCurl")
```

####1.Reading Dataset
First I read the raw data of New York City 2010 census data. This data is provided by John Keefe on his personal blog in form of Google Fusion Table.
```{r}
#Read rawdata from Github
mydataURL <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYC_Blocks_2010CensusData_Plus_Precincts.csv")
rawdata <- read.csv(text = mydataURL, quote = "")
```
####Cleaning Dataset and Renaming Variables
* Variables that I took directly from this dataset:  
    + 1. __precinc__:   NYPD police precinct number  
    
* Variables without Hispanic population as a seperate one  
    + 2. __P0010003__: Population of Hispanic White  
    + 3. __P0010004__: Population of Hispanic Black  
    + 4. __P0010005__: Population of Hispanic American Indian and Alaska Native  
    + 5. __P0010006__: Population of Hispanic Asian  
    + 6. __P0010007__: Population of Native Hawaiian and Other Pacific Islander  
    + 7. __P0010008__: Population of Other Hispanic Races  
    + 8. __P0010009__: Population of Other Hispanic with Mixed Races  
    
* Variables with Hispanic population as a seperate one  
    + 9. __P0020001__:  Whole population  
    + 10. __P0020005__:  Population of White alone  
    + 11. __P0020002__:  Population of Hispanic  
    + 12. __P0020008__:  Population of Asian alone  
    + 13. __P0020006__:  Population of Black alone  
    + 14. __P0020007__:  Population of American Indian and Alaska Native   
    + 15. __P0020009__:  Population of Native Hawaiian and Other Pacific Islander  
    + 16. __P0020010__:  Population of Others alone  
    + 17. __P0020011__： Population of Mixed  

I took 17 out 135 variables from this raw dataset. First I deleted all of the columns that has no precinct information. In doing this, we are not losing informations because all of the columns without information have no census information either. I also sorted the whole dataset by precinct. Then I changed the original column names into their corresponding races.  
* Explaination of terms used in this datasets:  
    + __Precicnt_: Also known as police precinct; New York City is divided into 76 geographical areas called precincts.
```{r}
#Delete the row with "Princinct = NA"
rawdata <- rawdata[complete.cases(rawdata$precinct),]
#Arrange the rawdata in ascending order with column "Precinct"
rawdata <- arrange(rawdata,precinct)
#Change the column names into corresponding races
censusdata <- rename(rawdata, Population = P0020001, HWhite = P0010003, HBlack = P0010004, HAsian = P0010006, HNative  = P0010005, HNativepac  = P0010007, HOthers = P0010008, HMixed = P0010009, White = P0020005, Hispanic = P0020002, Asian = P0020008, Black = P0020006, Native = P0020007, Nativepac = P0020009, Others = P0020010, Mixed = P0020011)
```


####2. Creating a new dataframe 
Here I created a new dataframe called "censusstat" to store the information we need from the raw dataset. I took the existing precinct from the raw dataset and made it into a new column in "censusstat" dataframe without repetitive precinct numbers so that we can sum up the information of different census blocks that are from the same precinct.

```{r}
#Track the existing precinct data in the raw data set
existingprecinct <- distinct(select(censusdata,precinct))
#Record the total number of precinct
precinctNumber <- length(existingprecinct) 
#Create a new data frame to store cleaned data
censusstat <- data.frame("Precinct" = existingprecinct)
#Arrange the censusstat in ascending order with column "Precinct"
censusstat <- arrange(censusstat,precinct)
```

####3.Calculating Population
In this chunck, I basically sum up the population by race and ethnicity (Hispanic only) and store them in "censusstat" in corresponding to the precinct.  

```{r}
#Calculate the population for each race without incorperating hispanic. Take white as an example, the numbers in each column represent the population of non-hispanic white.
#Since the data type of each column from the rawdata set is list instead of vector, I created a transitive dataframe called "popstat" to help me to store the result of each summation temporarily. 
censusdata_df = tbl_df(censusdata)
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Population = sum(Population))
censusstat$Population <- unlist(popstat[2]) #transfer a list into a vector
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(White = sum(White))
censusstat$White <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Asian = sum(Asian))
censusstat$Asian <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Black = sum(Black))
censusstat$Black <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Hispanic= sum(Hispanic))
censusstat$Hispanic<- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Native = sum(Native))
censusstat$Native<- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Nativepac = sum(Nativepac))
censusstat$Nativepac <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Others = sum(Others))
censusstat$Others <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(Mixed = sum(Mixed))
censusstat$Mixed <- unlist(popstat[2])
#Since "Native Indians"", "Native Islanders"" and "Others" have very small population, 
#I combined these three catogries into one variable and call it "NativeOther".
censusstat <- censusstat%>% mutate(NatvieOther =  Native + Nativepac + Others)

#Calculate the population for each race incorperating hispanic. Again, the numbers in each column for white combine white and hispanic white.
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HWhite = sum(HWhite))
censusstat$HWhite <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HAsian = sum(HAsian))
censusstat$HAsian <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HBlack = sum(HBlack))
censusstat$HBlack <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HNative = sum(HNative))
censusstat$HNative<- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HNativepac = sum(HNativepac))
censusstat$HNativepac <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HOthers = sum(HOthers))
censusstat$HOthers <- unlist(popstat[2])
popstat <- censusdata_df %>% group_by(precinct) %>% summarise(HMixed = sum(HMixed))
censusstat$HMixed <- unlist(popstat[2])

#Take the difference between hispanic and non-hispanic incorporated population to get the population of hispanic for each race.
censusstat$HisWhite = censusstat$HWhite - censusstat$White
censusstat$HisAsian = censusstat$HAsian - censusstat$Asian
censusstat$HisBlack = censusstat$HBlack - censusstat$Black
censusstat$HisNative = censusstat$HNative - censusstat$Native
censusstat$HisNativepac = censusstat$HNativepac - censusstat$Nativepac
censusstat$HisOthers = censusstat$HOthers - censusstat$Others
censusstat$HisMixed = censusstat$HMixed - censusstat$Mixed
censusstat <- censusstat%>% mutate(HisNatvieOther =  HisNative + HisNativepac + HisOthers)

#Delete the orginal variables "Native Indians"", "Native Islanders"" and "Others".
censusstat <- censusstat[-c(7:9, 12:18, 22:24)]
censusstat <- rename(censusstat, Precinct = precinct)
```

4. Remove unnecessary data to free up memory and view the cleaned dataset.
```{r}
#remove unnecessary data
rm(mydataURL, rawdata, existingprecinct, precinctNumber, censusdata, popstat, censusdata_df)
View(censusstat)
#setwd("/Users/Pinnzhang/Desktop")
write.csv(censusstat, file = "NYC_2010_Census_Data_Cleaned.csv")
```
*The original dataset is downloaded from URL:
    https://www.google.com/fusiontables/DataSource?dsrcid=767562#rows:id=1
 The documentation of this dataset is accessible from URL(Page 64): 
    https://www.documentcloud.org/documents/87708-pl94-171-1.html#document/p64


In order to make this RMD file  reproducable, I downloaded the file and pushed it onto my github with an open access.
    Github link for this dataset URL:
    https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYC_Blocks_2010CensusData_Plus_Precincts.csv
