---
title: "Clean St. Louis Schools Academic Performances"
author: "Ibuki Ogasawara"
date: "February 20, 2016"
output: html_document
---
This R markdown file cleans the datasets of academic performance in St. Louis public schools in 2015. Before running the code, some initial steps are required. (See footnotes.)  

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #summarize, arrange, select
```

####1. Download and Read Datasets from GitHub
```{r}
mydata <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED%20Academic%20Performance/SLED_Acad_Perf.csv")
SLED_ACAD_PERF <- read.csv(text = mydata, quote = "")

#Check the names of the columns and renames them
names(SLED_ACAD_PERF)
names(SLED_ACAD_PERF) <- c("X", "School", "Subject", "Category", "Type", "GradedSub", "Year", "Accountable", "Participant", "Reportable", "LevelUnknown", "LevelUnknown_Pct", "BelowBasic", "BelowBasic_Pct", "Basic", "Basic_Pct", "Proficient", "Proficient_Pct", "Advanced", "Advanced_Pct", "MapIndex", "MeanScore", "MedianTerranova")
```

####2. Select Columns
We will first select only the columns that we use. Namely: "School Name", "Content Area", "Accountable", "Reportable", "Below Basic", "Basic", "Proficient", "Advanced", "Mean Scale Score". 
```{r}
SLED_ACAD_Selected <- dplyr::select(SLED_ACAD_PERF, School, Subject, Accountable, Reportable, BelowBasic, Basic, Proficient, Advanced, MeanScore)
```

#####3. Filter the Data Frame
We separate the data frames of overall resutls and of each school.
```{r}
#Overall results
SLED_ACAD_Filtered1 <- filter(SLED_ACAD_Selected, School == '"State Overall"' | School == '"District Overall"')
#Results by school
SLED_ACAD_Filtered2 <- slice(SLED_ACAD_Selected, 35:488)
```

####4. Assign School Numbers
We will assign school number to each school by extracting the nubmers in the parentheses in the "School" column.
```{r}
#Function that extracts the 6th last character through 3rd last character of characters. 
substrRight_2 <- function(x){
  substr(x, nchar(x)-5, nchar(x)-2)
}
#Apply the function and convert the output into numeric
SLED_ACAD_Filtered2$SchNum <- as.numeric(substrRight_2(as.character(SLED_ACAD_Filtered2$School)))
```

####5. Replace NA with 0
Now, we replace "NA" with 0 so that we can carry out computations later.
```{r}
SLED_ACAD_Filtered2[is.na(SLED_ACAD_Filtered2)] <- 0
```

####6. Reorder and Select the Data Frame
- Bring the "SchNum" column to next to the "School"  
* We also delete some rows(schools) that has only 0 in the cells.  
        - 40: Beaumont CTE (1250)  
        - 62-64: Carnahan Sch (1500)  
        - 182-184: Gateway High (1220)  
        - 378-380: Roosvelt High (1680)  
        - 413:417: Vashon High (1830) and Sumner High (1800)
```{r}
SLED_ACAD_Reordered <- SLED_ACAD_Filtered2[, c("School", "SchNum", "Subject", "Accountable", "Reportable", "BelowBasic", "Basic", "Proficient", "Advanced", "MeanScore")]
SLED_ACAD_Filtered3 <- SLED_ACAD_Reordered[-c(40, c(62:64), c(182:184), c(378:380), c(413:417)),]
```

####7. Compute Number of Students by Category
Compute the means of the number of studnets in each category by school and subject.
```{r}
#Columns we want to group by
grp_cols <- names(SLED_ACAD_Filtered3)[c(2, 3)] #Select "School" and "Subject"
#Convert character vector to list of symbols
dots <- lapply(grp_cols, as.symbol)

Accountable_counts <- SLED_ACAD_Filtered3 %>%
        group_by_(.dots=dots) %>%
        summarise("Accountable" = sum(Accountable))
Reportable_counts <- SLED_ACAD_Filtered3 %>%
        group_by_(.dots=dots) %>%
        summarise("Reportable" = sum(Reportable))
BelowBasic_counts <- SLED_ACAD_Filtered3 %>%
        group_by_(.dots=dots) %>%
        summarise("BelowBasic" = sum(BelowBasic))
Basic_counts <- SLED_ACAD_Filtered3 %>%
        group_by_(.dots=dots) %>%
        summarise("Basic" = sum(Basic))
Proficient_counts <- SLED_ACAD_Filtered3 %>%
        group_by_(.dots=dots) %>%
        summarise("Proficient" = sum(Proficient))
Advanced_counts <- SLED_ACAD_Filtered3 %>% 
        group_by_(.dots=dots) %>%
        summarise("Advanced" = sum(Advanced))
MeanScore_sum <- SLED_ACAD_Filtered3 %>% 
        group_by_(.dots=dots) %>%
        summarise("MeanScore" = sum(MeanScore))
```

####8. Filter the Data Frame by Subject
Now, we merge them and separate the table into "English", "Math", and "Science" and merge them again into one data frame at the end. 
```{r}
#Check if the tables have the same dimensions.
sapply(list(Accountable_counts, Reportable_counts, BelowBasic_counts, Basic_counts, Proficient_counts, Advanced_counts, MeanScore_sum), dim)

Merged_table <- cbind.data.frame(Accountable_counts, Reportable_counts[, 3], BelowBasic_counts[, 3], Basic_counts[, 3], Proficient_counts[, 3], Advanced_counts[, 3], MeanScore_sum[, 3])

English <- filter(Merged_table, Subject=='"Eng. Language Arts"')
Math <- filter(Merged_table, Subject=='"Mathematics"')
Science <- filter(Merged_table, Subject=='"Science"')
```

####8.1. Rename the Data Frames Filtered by Subject
Now, we will rename the columns with certain abbreviations
Eng:English / Mat:Mathematics / Sci:Science
Ac: Acountable / Re: Reportable / BB: BelowBasic / Ba: Basic / Pr:Proficient / Ad:Advanced / MS:MeanScore
```{r}
names(English) <- c("SchNum", "Subject", "EngAc", "EngRe",
                    "EngBB", "EngBa", "EngPr", "EngAd", "EngMS")
names(Math) <- c("SchNum", "Subject", "MatAc", "MatRe",
                 "MatBB", "MatBa", "MatPr", "MatAd", "MatMS")
names(Science) <- c("SchNum", "Subject", "SciAc", "SciRe",
                    "SciBB", "SciBa", "SciPr", "SciAd", "SciMS")
```

####8.2 Merge the Data Frames
Finally, we will merge them into one table. 
```{r}
SLED_ACAD_Cleaned <- cbind.data.frame(English[, -2], Math[, 3:9], Science[, 3:9])
```

####9. Save the Data Frame
```{r}
write.csv(SLED_ACAD_Cleaned, "SLED_Academic_Peformance_2015_Cleaned.csv")
```

#### Footnote:

The original data source: Missouri Department of Elementary & Secondary Education   
http://mcds.dese.mo.gov/guidedinquiry/Achievement%20Level%20%204%20Levels/Achievement%20Level%204%20Report%20-%20Public.aspx?rp:DistrictCode=048914  

You can choose "School Year," "Content Area," "Category," etc that you are interested in.   
The dataset has information about the academic performance of all the public schools in St. Louis City. 