---
title: "Clean St. Louis Public Schools Demographics"
author: "Ibuki Ogasawara"
date: "February 17, 2016"
output: html_document
---
This R markdown file cleans the datasets of demographics in St. Louis public school from 2008 to 2015. Before running the code, some initial steps are required. (See footnotes.)  

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #mutate, arrange, select
```

####1. Download and Read Datasets from GitHub
```{r}
mydata <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/SLED_Demographics/SLED_Demographic_Data.csv")
SLED_DEMO <- read.csv(text = mydata, quote = "")
```

####2. Compute Populations by Race
First, we will add the columns of counts of students in each ethnic groups by multiplying "total" and each percentage. 
```{r}
SLED_DEMO_Counts <- mutate(SLED_DEMO, 
                    Asian = round(SLED_DEMO$TOTAL*SLED_DEMO$Asian_Per/100, digits=0), 
                    Black = round(SLED_DEMO$TOTAL*SLED_DEMO$Black_Per/100, digits=0),
                    Hispanic = round(SLED_DEMO$TOTAL*SLED_DEMO$Hispanic_Per/100, digits=0),
                    Indian = round(SLED_DEMO$TOTAL*SLED_DEMO$Indian_Per/100, digits=0),
                    White = round(SLED_DEMO$TOTAL*SLED_DEMO$White_Per/100, digits=0))
```

####2.1 Compute the Number of Free Reduced Lunch
Get rid of "*" from the "FreeReducedLunch_FTE_Per" column. And compute the total counts.
```{r}
SLED_DEMO_Counts$FreeReducedLunch_FTE_Per[SLED_DEMO_Counts$FreeReducedLunch_FTE_Per=="*"] <- NA
FTE_per_cha <- as.character(SLED_DEMO_Counts$FreeReducedLunch_FTE_Per)
FTE_per_num <- as.numeric(FTE_per_cha)
SLED_DEMO_Counts2 <- mutate(SLED_DEMO_Counts,
                            FTE = round(SLED_DEMO_Counts$TOTAL*FTE_per_num/100,
                                        digits=0))
```

####3. Assign School Number
We will assign school number to each school by extracting the nubmers in the parentheses in the SCHOOL_NAME column. 
```{r}
#Function that extracts the 5th last character through 2rd last character of characters. 
substrRight_1 <- function(x){
  substr(x, nchar(x)-4, nchar(x)-1)
}
#Apply the function and convert the output into numeric
SLED_DEMO_Counts2$SchNum <- as.numeric(substrRight_1(as.character(SLED_DEMO_Counts2$SCHOOL_NAME)))
```

####4. Select Columns
Lastly, we will only select the columns that we created and arrange by year and school number. 
```{r}
SLED_DEMO_Selected <- dplyr::select(SLED_DEMO_Counts2,SCHOOL_NAME:TOTAL, Asian:SchNum)
#Move two of the columns "SchNum" and "TOTAL" so that the table makes more sense. 
SLED_DEMO_Reordered <- SLED_DEMO_Selected[, c("SCHOOL_NAME","YEAR","SchNum","Asian","Black","Hispanic","Indian","White","FTE","TOTAL")] 
#Finally, we will arrange the rows and rename columns. 
SLED_DEMO_Arranged <- arrange(SLED_DEMO_Reordered, YEAR, SchNum)
names(SLED_DEMO_Arranged) <- c("School","Year","SchNum","Asian","Black","Hispanic","Indian","White","FTE","Total")
```

####9. Save the data frame
```{r}
write.csv(SLED_DEMO_Arranged, "SLED_Demographic_Cleaned.csv")
```

#### Footnote: 

The original data source: Missouri Department of Elementary & Secondary Education  
http://mcds.dese.mo.gov/quickfacts/SitePages/DistrictInfo.aspx?ID=__bk8100130013005300130013005300  

Under "ST. LOUIS CITY -- Resources and Process", "Student Demographics" - "Building"  

Initial cleaning process:  
1. In Excel, deleted commas in one the school names (EARL NANCE, SR. ELEM. (5610))  
2. In Excel, deleted two columns in which all the entries were 0s.  
3. Import the dataset into Rstudio and deleted commas in the column "TOTAL" by
```{r, eval=FALSE}
total <- SLED.Demographic.Data$TOTAL
total.cha <- as.character(total)
total.num <- as.numeric(gsub(",", "", total.cha))
SLED.Demographic.Data$TOTAL <- total.num
write.csv(SLED.Demographic.Data, "SLED_Demographic_Data.csv")
```

