---
title: "Clean New York City Public Schools Graduation Rate"
author: "Ibuki Ogasawara"
date: "March 10, 2016"
output: html_document
---
This R markdown file cleans the datasets of demographics of New York City public schools from 2011 to 2015. Before running the code, some initial steps are required. (See footnotes.) 

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #summarize, arrange, select, filter
```


####1. Download and Read Dataset from GitHub 
```{r message=FALSE}
mydata <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYC%20Graduation%20Rate/NYED_Grad_Results.csv")
NYED_GRAD <- read.csv(text = mydata, quote = "")
```

####2. Reduce Columns
Omit all the "percentage" columns. 
```{r message=FALSE}
#Rename columns
names(NYED_GRAD)
names(NYED_GRAD) <- c("X", "DBN", "School", "CohortYear", "Category", "Total", "Total_Grad", "Total_Grad_Pct", "All_Regents", "All_Regents_Pct_C", "All_Regents_Pct_G", "Adv_Regents", "Adv_Regents_Pct_C", "Adv_Regents_Pct_G", "Regents_wo_Advanced", "Regents_wo_Advanced_Pct_C", "Regents_wo_Advanced_Pct_G", "Local", "Local_Pct_C", "Local_Pct_G", "Still_Enroll","Still_Enroll_Pct", "DropOut", "DropOut_Pct")

NYED_GRAD_Reduced_columns <- select(NYED_GRAD, DBN, School, CohortYear, Category, Total, Total_Grad, All_Regents, Adv_Regents, Local, Still_Enroll, DropOut)
```

####3. Delete August Graduation Rows
```{r}
#We will delete "August" graduation rows and keep the ones for "June"
table(NYED_GRAD$Category)
NYED_GRAD_Reduced <- filter(NYED_GRAD_Reduced_columns, !(Category=='"4 Year August"'| Category=='"5 Year August"'))
table(NYED_GRAD_Reduced$Category)
```

####4.0. Keep 4 and 6 Year Graduation
We are going to keep only 4 and 6 year graduation results. To do this, we create a simplified column with 4, 5, 6 and filter 4 and 6. 
```{r}
NYED_GRAD_Reduced$Category <- substr(NYED_GRAD_Reduced$Category, 2, 2) #The second letter of '"4 year ~"'
#filter
NYED_GRAD_46 <- filter(NYED_GRAD_Reduced, Category=="4"|Category=="6")
```

####4.1. Keep 4 and 6 Year Graduation
Create a graduation year column
```{r}
NYED_GRAD_46$Grad_Year <- as.numeric(as.character(NYED_GRAD_46$CohortYear))+as.numeric(as.character(NYED_GRAD_46$Category))
#Remove double quotations from "DBN" and "School" columns.
substr_middle <- function(x){
  substr(x, 2, nchar(x)-1)
}
NYED_GRAD_46$DBN <- substr_middle(as.character(NYED_GRAD_46$DBN))
NYED_GRAD_46$School <- substr_middle(as.character(NYED_GRAD_46$School))
```

####5. Filter by Year
Now we separate the dataset by each graduation year form 2006 to 2015.
```{r}
#2006
NYED_GRAD_06 <- filter(NYED_GRAD_46, Grad_Year=="2006")
#2006 only has 4 year graduation. 
#Rename and reorder the columns
names(NYED_GRAD_06)
names(NYED_GRAD_06) <- c("DBN","School","CohortYear","Category","Total4", "TotalGrad4","AllRegents4","AdvRegents4","Local4", "StillEnroll4","DropOut4","Year")
NYED_GRAD_06_Selected <- select(NYED_GRAD_06, DBN, School, Year, Total4, TotalGrad4, AllRegents4, AdvRegents4, Local4, StillEnroll4, DropOut4)
```

####6. Check the Class of Variables
Check if we need to change the class of the columns. In this case, they are all numeric. 
```{r}
list_of_columns <- list(NYED_GRAD_06_Selected$DBN, NYED_GRAD_06_Selected$School, NYED_GRAD_06_Selected$Year, NYED_GRAD_06_Selected$Total4, NYED_GRAD_06_Selected$TotalGrad4, NYED_GRAD_06_Selected$AllRegents4, NYED_GRAD_06_Selected$AdvRegents4, NYED_GRAD_06_Selected$Local4, NYED_GRAD_06_Selected$StillEnroll4, NYED_GRAD_06_Selected$DropOut4)
sapply(list_of_columns, class)
```

####7. Save the Data Frame
```{r}
write.csv(NYED_GRAD_06_Selected, "NYED_Graduation_Rate_2006_Cleaned.csv")
```

####8. Repeat the Process
From 2005 (Graduation year of 2010), there are June and August. We will only use the one for June. 
```{r}
#2015
NYED_GRAD_15 <- filter(NYED_GRAD_46, Grad_Year=="2015")
#From 2007 there are 4 and 6 year graduations. 
NYED_GRAD_15_4 <- filter(NYED_GRAD_15, Category=="4")
NYED_GRAD_15_6 <- filter(NYED_GRAD_15, Category=="6")
#Rename and reorder the columns of the two dataframes.
names(NYED_GRAD_15)
names(NYED_GRAD_15_4) <- c("DBN","School","CohortYear","Category","Total4","TotalGrad4","AllRegents4","AdvRegents4","Local4","StillEnroll4","DropOut4","Year")
names(NYED_GRAD_15_6) <- c("DBN","School","CohortYear","Category","Total6","TotalGrad6","AllRegents6","AdvRegents6","Local6","StillEnroll6","DropOut6","Year")

NYED_GRAD_15_4_Selected <- select(NYED_GRAD_15_4, DBN, School, Year, Total4, TotalGrad4,AllRegents4, AdvRegents4, Local4, StillEnroll4, DropOut4)
NYED_GRAD_15_6_Selected <- select(NYED_GRAD_15_6, DBN, Total6, TotalGrad6,AllRegents6, AdvRegents6, Local6, StillEnroll6, DropOut6)
NYED_GRAD_15_Joined <- full_join(NYED_GRAD_15_4_Selected, NYED_GRAD_15_6_Selected, by="DBN")
```
Check if we need to change the class of the columns. 
```{r}
list_of_columns <- list(NYED_GRAD_15_Joined$DBN, NYED_GRAD_15_Joined$School, NYED_GRAD_15_Joined$Year, NYED_GRAD_15_Joined$Total4, NYED_GRAD_15_Joined$TotalGrad4, NYED_GRAD_15_Joined$AllRegents4, NYED_GRAD_15_Joined$AdvRegents4, NYED_GRAD_15_Joined$Local4, NYED_GRAD_15_Joined$StillEnroll4, NYED_GRAD_15_Joined$DropOut4)
sapply(list_of_columns, class)
```
Finish
```{r}
write.csv(NYED_GRAD_15_Joined, "NYED_Graduation_Rate_2015_Cleaned.csv")
```

#### Footnote:

The original dataset is downloaded from the website:  
http://schools.nyc.gov/Accountability/data/GraduationDropoutReports/default.htm

The dataset has information about the Mathematics test results of all the public schools in New York City from 3rd to 8th grade.  
The excel file has several sheets and we will only use "All Student" spread sheet.  
Initial clean: Deleting commas in all "_Num" columns in R with the following code.
```{r eval=FALSE}
NYC_Grad <- read.csv("~/Desktop/NYC_Graducation_Resuts.csv")

#Total Cohort
total <- NYC_Grad$Total_Cohort_Num
total.num <- as.numeric(gsub(",", "", as.character(total)))
NYC_Grad$Total_Cohort_Num <- total.num

#Total Graduation
grad <- NYC_Grad$Total_Grads_Num
grad.num <- as.numeric(gsub(",", "", as.character(grad)))
NYC_Grad$Total_Grads_Num <- grad.num

#Total Regents
regents <- NYC_Grad$Total_Regents_Num
regents.num <- as.numeric(gsub(",", "", as.character(regents)))
NYC_Grad$Total_Regents_Num <- regents.num

#Advanced Regents
adregents <- NYC_Grad$Advanced_Regents_Num
adregents.num <- as.numeric(gsub(",", "", as.character(adregents)))
NYC_Grad$Advanced_Regents_Num <- adregents.num

#Regents without Advanced
nadregents <- NYC_Grad$Regents_wo_Advanced_Num
nadregents.num <- as.numeric(gsub(",", "", as.character(nadregents)))
NYC_Grad$Regents_wo_Advanced_Num <- nadregents.num

#Local
local <- NYC_Grad$Local_Num
local.num <- as.numeric(gsub(",", "", as.character(local)))
NYC_Grad$Local_Num <- local.num

#Still Enrolled
enroll <- NYC_Grad$Still_Enrolled_Num
enroll.num <- as.numeric(gsub(",", "", as.character(enroll)))
NYC_Grad$Still_Enrolled_Num <- enroll.num

#Dropped Out
drop <- NYC_Grad$Dropped_Out_Num
drop.num <- as.numeric(gsub(",", "", as.character(drop)))
NYC_Grad$Dropped_Out_Num <- drop.num


write.csv(NYC_Grad, "NYED_Grad_Results.csv")
```