---
title: "Clean New York City Public Schools ELA Performance"
author: "Ibuki Ogasawara"
date: "March 6, 2016"
output: html_document
---
This R markdown file cleans the datasets of ELA test performance of New York City public schools from 2006 to 2015. Before running the code, some initial steps are required. (See footnotes.) 

####0. Install Required Packages
```{r message=FALSE}
if(!require(RCurl))install.packages("RCurl")
library(RCurl) #getURL(download datasets from a website)
if(!require(dplyr))install.packages("dplyr")
library(dplyr) #summarize, arrange, select, filter
```

####1. Download and Read Datasets from GitHub
```{r message=FALSE}
mydata1 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYC%20Academic%20Perfomance/NYC_ELA_06-12.csv")
NYED_ELA_PERF_0612 <- read.csv(text = mydata1, quote = "")

mydata2 <- getURL("https://raw.githubusercontent.com/pinnnnnn/MAP_Spring_2016/master/NYC%20Academic%20Perfomance/NYC_ELA_13-15.csv")
NYED_ELA_PERF_1315 <- read.csv(text = mydata2, quote = "")

names(NYED_ELA_PERF_0612)
names(NYED_ELA_PERF_0612) <- c("DBN", "Grade", "Year", "Category", "NumTest", "MeanScore", "L1Num", "L1Pct", "L2Num", "L2Pct","L3Num", "L3Pct","L4Num", "L4Pct","L34Num", "L34Pct")
names(NYED_ELA_PERF_1315)
names(NYED_ELA_PERF_1315) <- c("DBN", "School", "Grade", "Year", "Category", "NumTest", "MeanScore", "L1Num", "L1Pct", "L2Num", "L2Pct","L3Num", "L3Pct","L4Num", "L4Pct","L34Num", "L34Pct")
```

####2. Select Columns
```{r}
#Select the columns that we will use. 
NYED_ELA_PERF_0612 <- select(NYED_ELA_PERF_0612, DBN, Grade, Year, NumTest, MeanScore, L1Num, L2Num, L3Num, L4Num, L34Num, L34Pct) #"Category" and all the percentages except for Level3+4 are excluded
NYED_ELA_PERF_1315 <- select(NYED_ELA_PERF_1315, DBN, School, Grade, Year, NumTest, MeanScore, L1Num, L2Num, L3Num, L4Num, L34Num, L34Pct) #"Category", and all the percentages except for Level3+4 are excluded
```

####3. Filter by Year
Now, we will divide the dataset by years. 
```{r}
ELA_2006 <- filter(NYED_ELA_PERF_0612, Year=="2006")
ELA_2007 <- filter(NYED_ELA_PERF_0612, Year=="2007")
ELA_2008 <- filter(NYED_ELA_PERF_0612, Year=="2008")
ELA_2009 <- filter(NYED_ELA_PERF_0612, Year=="2009")
ELA_2010 <- filter(NYED_ELA_PERF_0612, Year=="2010")
ELA_2011 <- filter(NYED_ELA_PERF_0612, Year=="2011")
ELA_2012 <- filter(NYED_ELA_PERF_0612, Year=="2012")
ELA_2013 <- filter(NYED_ELA_PERF_1315, Year=="2013")
ELA_2014 <- filter(NYED_ELA_PERF_1315, Year=="2014")
ELA_2015 <- filter(NYED_ELA_PERF_1315, Year=="2015")
```

####4. Compute Mean Score and Improvement
We will use the rows of "5"th grade and "All Grades" to measuare the performance of each school.
From 2006 to 2010, the "MeanScore"s for "All Grades" are missing. So use "group_by" to compute the the values. 
Also, we will compute the "improvement" by taking the differences in the percentages of proficient students in comparison to 2006.  
```{r}
#2006
#Exclude the rows with "All Grades"
ELA_06_Filtered <- filter(ELA_2006, Grade=="3"|Grade=="4"|Grade=="5"|Grade=="6"|Grade=="7"|Grade=="8") 
#First, we compute "MeanScore" for "All Grades"
#convert "factor" into "numeric"
ELA_06_Filtered$MeanScore <- as.numeric(as.character(ELA_06_Filtered$MeanScore))
#Since "MeanScore" includes the character "s"s, "as.numeric" will produce "NA"s. 
MS_06 <- ELA_06_Filtered %>%
        group_by(DBN) %>%
        summarise("MS" = round(mean(MeanScore, na.rm=TRUE), digits=0))
ELA_06_AllG <- filter(ELA_2006, Grade=="All Grades")
#Arrange the datasets by DBN just in case
ELA_06_AllG <- arrange(ELA_06_AllG, DBN)
MS_06 <- arrange(MS_06, DBN)
#Add MS_06 to ELA_06_AllG
ELA_06_AllG$MeanScore <- MS_06[, 2]
names(ELA_06_AllG)[5] <- "MeanScore"
dim(ELA_06_AllG)
#Now, compute the "improvements"
ELA_06_AllG$Improv_AllG <- as.numeric(as.character(ELA_06_AllG$L34Pct))-as.numeric(as.character(ELA_06_AllG$L34Pct))
```

####5. Filter 5th Grade
```{r}
#Now, create the columns of 5th grade results
ELA_06_5th_Filtered <- filter(ELA_2006, Grade=="5")
ELA_06_5th_Arranged <- arrange(ELA_06_5th_Filtered, DBN)
dim(ELA_06_5th_Arranged)
#Get rid of unnecessary columns.
ELA_06_5th_Reduced <- ELA_06_5th_Arranged[, -c(2,3)] #"Grade" and "Year" columns. 
#Now, compute the "improvements"
ELA_06_5th_Reduced$Improv5th <- as.numeric(as.character(ELA_06_5th_Reduced$L34Pct)) - as.numeric(as.character(ELA_06_5th_Reduced$L34Pct))
```

####6. Join All Grade and 5th Grade Results
```{r}
#Change names of the columns for the join. 
names(ELA_06_5th_Reduced) <- c("DBN", "NumTest5th", "MeanScore5th", "L1Num5th", "L2Num5th", "L3Num5th", "L4Num5th", "L34Num5th", "L34Pct5th", "Improv5th")
ELA_06_AllG_5th <- merge(ELA_06_AllG, ELA_06_5th_Reduced, by="DBN", all=TRUE)

#Finally, we remove "Grade" and "Year" columns, and reorder/renames the columns. 
ELA_06_AllG_5th_Selected <- select(ELA_06_AllG_5th, DBN, NumTest, NumTest5th, MeanScore, MeanScore5th, L1Num, L1Num5th, L2Num, L2Num5th, L3Num, L3Num5th, L4Num, L4Num5th, L34Num, L34Num5th, L34Pct, L34Pct5th, Improv_AllG, Improv5th)
names(ELA_06_AllG_5th_Selected) <- c("DBN", "NumTest", "NumTest5th", "MeanScore", "MeanScore5th", "L1Num", "L1Num5th", "L2Num", "L2Num5th", "L3Num", "L3Num5th", "L4Num", "L4Num5th", "L34Num", "L34Num5th", "L34Pct", "L34Pct5th", "Improv", "Improv5th")
dim(ELA_06_AllG_5th_Selected)
```

####7. Check the Class of Variables
Check if we need to change the class of the columns. 
```{r}
list_of_columns <- list(ELA_06_AllG_5th_Selected$DBN, ELA_06_AllG_5th_Selected$NumTest, ELA_06_AllG_5th_Selected$NumTest5th, ELA_06_AllG_5th_Selected$MeanScore, ELA_06_AllG_5th_Selected$MeanScore_5, ELA_06_AllG_5th_Selected$L1Num, ELA_06_AllG_5th_Selected$L1Num_5, ELA_06_AllG_5th_Selected$L2Num, ELA_06_AllG_5th_Selected$L2Num_5, ELA_06_AllG_5th_Selected$L3Num, ELA_06_AllG_5th_Selected$L3Num_5, ELA_06_AllG_5th_Selected$L4Num, ELA_06_AllG_5th_Selected$L4Num_5, ELA_06_AllG_5th_Selected$L34Num, ELA_06_AllG_5th_Selected$L34Num_5, ELA_06_AllG_5th_Selected$L34Pct, ELA_06_AllG_5th_Selected$L34Pct_5, ELA_06_AllG_5th_Selected$Improv, ELA_06_AllG_5th_Selected$Improv_5)
sapply(list_of_columns, class)
```

####8. Change the Class to Numeric
Every column but "DBN" has to be numeric or integer. 
```{r}
ELA_06_AllG_5th_Selected$MeanScore <- ELA_06_AllG_5th_Selected$MeanScore$MS #This column was in fact a data frame. 
ELA_06_AllG_5th_Selected$MeanScore5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$MeanScore5th))[ELA_06_AllG_5th_Selected$MeanScore5th]
ELA_06_AllG_5th_Selected$L1Num <- as.numeric(levels(ELA_06_AllG_5th_Selected$L1Num))[ELA_06_AllG_5th_Selected$L1Num]
ELA_06_AllG_5th_Selected$L1Num5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L1Num5th))[ELA_06_AllG_5th_Selected$L1Num5th]
ELA_06_AllG_5th_Selected$L2Num <- as.numeric(levels(ELA_06_AllG_5th_Selected$L2Num))[ELA_06_AllG_5th_Selected$L2Num]
ELA_06_AllG_5th_Selected$L2Num5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L2Num5th))[ELA_06_AllG_5th_Selected$L2Num5th]
ELA_06_AllG_5th_Selected$L3Num <- as.numeric(levels(ELA_06_AllG_5th_Selected$L3Num))[ELA_06_AllG_5th_Selected$L3Num]
ELA_06_AllG_5th_Selected$L3Num5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L3Num5th))[ELA_06_AllG_5th_Selected$L3Num5th]
ELA_06_AllG_5th_Selected$L4Num <- as.numeric(levels(ELA_06_AllG_5th_Selected$L4Num))[ELA_06_AllG_5th_Selected$L4Num]
ELA_06_AllG_5th_Selected$L4Num5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L4Num5th))[ELA_06_AllG_5th_Selected$L4Num5th]
ELA_06_AllG_5th_Selected$L34Num <- as.numeric(levels(ELA_06_AllG_5th_Selected$L34Num))[ELA_06_AllG_5th_Selected$L34Num]
ELA_06_AllG_5th_Selected$L34Num5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L34Num5th))[ELA_06_AllG_5th_Selected$L34Num5th]
ELA_06_AllG_5th_Selected$L34Pct <- as.numeric(levels(ELA_06_AllG_5th_Selected$L34Pct))[ELA_06_AllG_5th_Selected$L34Pct]
ELA_06_AllG_5th_Selected$L34Pct5th <- as.numeric(levels(ELA_06_AllG_5th_Selected$L34Pct5th))[ELA_06_AllG_5th_Selected$L34Pct5th]
```
Check it again. 
```{r}
list_of_columns <- list(ELA_06_AllG_5th_Selected$DBN, ELA_06_AllG_5th_Selected$NumTest, ELA_06_AllG_5th_Selected$NumTest5th, ELA_06_AllG_5th_Selected$MeanScore, ELA_06_AllG_5th_Selected$MeanScore5th, ELA_06_AllG_5th_Selected$L1Num, ELA_06_AllG_5th_Selected$L1Num5th, ELA_06_AllG_5th_Selected$L2Num, ELA_06_AllG_5th_Selected$L2Num5th, ELA_06_AllG_5th_Selected$L3Num, ELA_06_AllG_5th_Selected$L3Num5th, ELA_06_AllG_5th_Selected$L4Num, ELA_06_AllG_5th_Selected$L4Num5th, ELA_06_AllG_5th_Selected$L34Num, ELA_06_AllG_5th_Selected$L34Num5th, ELA_06_AllG_5th_Selected$L34Pct, ELA_06_AllG_5th_Selected$L34Pct5th, ELA_06_AllG_5th_Selected$Improv, ELA_06_AllG_5th_Selected$Improv5th)
sapply(list_of_columns, class)
```

####9. Save the Data Frame 
```{r}
write.csv(ELA_06_AllG_5th_Selected, "NYED_ELA_2006_Cleaned.csv")
```

####10. Repeat the Process
From 2011, we do not need to compute the "MeanScore" for "All Grades"
```{r}
#2015
ELA_15_AllG <- filter(ELA_2015, Grade=="All Grades")
#Arrange the datasets by DBN just in case
ELA_15_AllG <- arrange(ELA_15_AllG, DBN)
dim(ELA_15_AllG)
#Now, compute the "improvements"
ELA_06_L34Pct <- select(ELA_06_AllG_5th_Selected, DBN, L34Pct)
names(ELA_06_L34Pct) <- c("DBN", "L34Pct_06")
ELA_15_AllG_Merged <- merge(ELA_15_AllG, ELA_06_L34Pct, by="DBN", all=TRUE)
ELA_15_AllG_Merged$Improv_AllG <- round(as.numeric(as.character(ELA_15_AllG_Merged$L34Pct))-as.numeric(as.character(ELA_15_AllG_Merged$L34Pct_06)), digits=2)

#Now, create the columns of 5th grade results
ELA_15_5th_Filtered <- filter(ELA_2015, Grade=="5")
ELA_15_5th_Arranged <- arrange(ELA_15_5th_Filtered, DBN)
dim(ELA_15_5th_Arranged)
#Get rid of unnecessary columns.
ELA_15_5th_Reduced <- ELA_15_5th_Arranged[, -c(3,4)] #"Grade" and "Year" columns. 
#Now, compute the "improvements"
ELA_06_L34Pct5th <- select(ELA_06_AllG_5th_Selected, DBN, L34Pct5th)
names(ELA_06_L34Pct5th) <- c("DBN", "L34Pct5th_06")
ELA_15_5th_Merged <- merge(ELA_15_5th_Reduced, ELA_06_L34Pct5th, by="DBN", all=TRUE)
ELA_15_5th_Merged$Improv5th <- round(as.numeric(as.character(ELA_15_5th_Merged$L34Pct))-as.numeric(as.character(ELA_15_5th_Merged$L34Pct5th_06)), digits=2)
#Change names of the columns for the join. 
names(ELA_15_5th_Merged) <- c("DBN", "School", "NumTest5th", "MeanScore5th", "L1Num5th", "L2Num5th", "L3Num5th", "L4Num5th", "L34Num5th", "L34Pct5th", "L34Pct5th_06", "Improv5th")
ELA_15_AllG_5th <- merge(ELA_15_AllG_Merged, ELA_15_5th_Merged, by="DBN", all=TRUE)

#Finally, we remove "Grade" and "Year" columns, and reorder/renames the columns. 
ELA_15_AllG_5th_Selected <- select(ELA_15_AllG_5th, DBN, School.x, NumTest, NumTest5th, MeanScore, MeanScore5th, L1Num, L1Num5th, L2Num, L2Num5th, L3Num, L3Num5th, L4Num, L4Num5th, L34Num, L34Num5th, L34Pct, L34Pct5th, Improv_AllG, Improv5th)
names(ELA_15_AllG_5th_Selected) <- c("DBN", "School", "NumTest", "NumTest5th", "MeanScore", "MeanScore5th", "L1Num", "L1Num5th", "L2Num", "L2Num5th", "L3Num", "L3Num5th", "L4Num", "L4Num5th", "L34Num", "L34Num5th", "L34Pct", "L34Pct5th", "Improv", "Improv5th")
dim(ELA_15_AllG_5th_Selected)
```
Check if we need to change the class of the columns. 
```{r}
list_of_columns <- list(ELA_15_AllG_5th_Selected$DBN, ELA_15_AllG_5th_Selected$School, ELA_15_AllG_5th_Selected$NumTest, ELA_15_AllG_5th_Selected$NumTest5th, ELA_15_AllG_5th_Selected$MeanScore, ELA_15_AllG_5th_Selected$MeanScore5th, ELA_15_AllG_5th_Selected$L1Num, ELA_15_AllG_5th_Selected$L1Num5th, ELA_15_AllG_5th_Selected$L2Num, ELA_15_AllG_5th_Selected$L2Num5th, ELA_15_AllG_5th_Selected$L3Num, ELA_15_AllG_5th_Selected$L3Num5th, ELA_15_AllG_5th_Selected$L4Num, ELA_15_AllG_5th_Selected$L4Num5th, ELA_15_AllG_5th_Selected$L34Num, ELA_15_AllG_5th_Selected$L34Num5th, ELA_15_AllG_5th_Selected$L34Pct, ELA_15_AllG_5th_Selected$L34Pct5th, ELA_15_AllG_5th_Selected$Improv, ELA_15_AllG_5th_Selected$Improv5th)
sapply(list_of_columns, class)
```
Every column but "DBN" has to be numeric or integer. 
```{r}
ELA_15_AllG_5th_Selected$MeanScore <- as.numeric(levels(ELA_15_AllG_5th_Selected$MeanScore))[ELA_15_AllG_5th_Selected$MeanScore]
ELA_15_AllG_5th_Selected$MeanScore5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$MeanScore5th))[ELA_15_AllG_5th_Selected$MeanScore5th]
ELA_15_AllG_5th_Selected$L1Num <- as.numeric(levels(ELA_15_AllG_5th_Selected$L1Num))[ELA_15_AllG_5th_Selected$L1Num]
ELA_15_AllG_5th_Selected$L1Num5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L1Num5th))[ELA_15_AllG_5th_Selected$L1Num5th]
ELA_15_AllG_5th_Selected$L2Num <- as.numeric(levels(ELA_15_AllG_5th_Selected$L2Num))[ELA_15_AllG_5th_Selected$L2Num]
ELA_15_AllG_5th_Selected$L2Num5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L2Num5th))[ELA_15_AllG_5th_Selected$L2Num5th]
ELA_15_AllG_5th_Selected$L3Num <- as.numeric(levels(ELA_15_AllG_5th_Selected$L3Num))[ELA_15_AllG_5th_Selected$L3Num]
ELA_15_AllG_5th_Selected$L3Num5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L3Num5th))[ELA_15_AllG_5th_Selected$L3Num5th]
ELA_15_AllG_5th_Selected$L4Num <- as.numeric(levels(ELA_15_AllG_5th_Selected$L4Num))[ELA_15_AllG_5th_Selected$L4Num]
ELA_15_AllG_5th_Selected$L4Num5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L4Num5th))[ELA_15_AllG_5th_Selected$L4Num5th]
ELA_15_AllG_5th_Selected$L34Num <- as.numeric(levels(ELA_15_AllG_5th_Selected$L34Num))[ELA_15_AllG_5th_Selected$L34Num]
ELA_15_AllG_5th_Selected$L34Num5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L34Num5th))[ELA_15_AllG_5th_Selected$L34Num5th]
ELA_15_AllG_5th_Selected$L34Pct <- as.numeric(levels(ELA_15_AllG_5th_Selected$L34Pct))[ELA_15_AllG_5th_Selected$L34Pct]
ELA_15_AllG_5th_Selected$L34Pct5th <- as.numeric(levels(ELA_15_AllG_5th_Selected$L34Pct5th))[ELA_15_AllG_5th_Selected$L34Pct5th]
```
Check it again. 
```{r}
list_of_columns <- list(ELA_15_AllG_5th_Selected$DBN, ELA_15_AllG_5th_Selected$School, ELA_15_AllG_5th_Selected$NumTest, ELA_15_AllG_5th_Selected$NumTest5th, ELA_15_AllG_5th_Selected$MeanScore, ELA_15_AllG_5th_Selected$MeanScore5th, ELA_15_AllG_5th_Selected$L1Num, ELA_15_AllG_5th_Selected$L1Num5th, ELA_15_AllG_5th_Selected$L2Num, ELA_15_AllG_5th_Selected$L2Num5th, ELA_15_AllG_5th_Selected$L3Num, ELA_15_AllG_5th_Selected$L3Num5th, ELA_15_AllG_5th_Selected$L4Num, ELA_15_AllG_5th_Selected$L4Num5th, ELA_15_AllG_5th_Selected$L34Num, ELA_15_AllG_5th_Selected$L34Num5th, ELA_15_AllG_5th_Selected$L34Pct, ELA_15_AllG_5th_Selected$L34Pct5th, ELA_15_AllG_5th_Selected$Improv, ELA_15_AllG_5th_Selected$Improv5th)
sapply(list_of_columns, class)
```
Done!
```{r}
write.csv(ELA_15_AllG_5th_Selected, "NYED_ELA_2015_Cleaned.csv")
#At the bottom of the datasets, there are about 100 school DBN rows that do not have any attributes. This is the result of merging "ELA_06" with "ELA_15" when calculating "Improvements". Obviously, some schools had been closed and others have been newly established over the years. 
```

#### Footnote:
The original dataset is downloaded from the website:  
http://schools.nyc.gov/Accountability/data/TestResults/ELAandMathTestResults  
and  
http://schools.nyc.gov/NR/exeres/05289E74-2D81-4CC0-81F6-E1143E28F4C4,frameless.htm  

The dataset has information about the English Language Arts (ELA) test results of all the public schools in New York City from 3rd to 8th grade.  
The excel file has several sheets and we will only use "All Student" spread sheet. 